### 2022-08-10 doc/src/VSLevel.tex
Replace "I/O MMU" with "IOMMU"

To ensure proper support for interrupts, two conditions must be met
before a hypervisor may allow a guest OS running in a virtual machine
to directly control a physical device that sends MSIs:
First, each virtual hart must have a guest interrupt file assigned to
it, giving each its own apparent IMSIC within the virtual machine.
Second, interrupts from the device must be signaled by wire through
an APLIC that can translate these interrupts into MSIs, or the
system must have an I/O MMU that can translate the addresses of MSI
memory writes made by the device itself.

If a guest OS directly controls a device capable of sending MSIs, it
will naturally configure MSIs at the device with the guest physical
addresses the OS sees for the IMSICs of its virtual harts, not knowing
that these addresses are only virtual.
When the device performs a memory write for an MSI, the destination
address of this write must be translated by the I/O MMU from the
guest physical address assigned by the guest OS into the true physical
address of the target guest interrupt file, using a translation table
supplied by the hypervisor.

By design, the translation an I/O MMU must do for device MSIs is
fundamentally no different than the address translation the I/O MMU
already must perform for other memory accesses from the same device,
converting guest physical addresses into true physical addresses.
Because each virtual hart is assigned a dedicated, physical guest
interrupt file that is indistinguishable from a true supervisor-level
interrupt file, no translation is needed for the data of an MSI write,
which specifies the interrupt's identity number in the target interrupt
file.

@@

\item
Modify the relevant translation tables at all I/O MMUs so that MSIs for
this virtual interrupt file are now sent to the new physical interrupt
file.
Likewise, if any interrupts at an APLIC are forwarded by MSIs to the old
interrupt file, reconfigure the APLIC to send them to the new interrupt
file.
As needed, synchronize with all I/O MMUs and APLICs to ensure that no
straggler MSIs will arrive at the old interrupt file after this step.
Synchronizing with an APLIC can be accomplished using the
algorithm of Section~\ref{sec:AdvPLIC-MSISync}.

>>>>

To ensure proper support for interrupts, two conditions must be met
before a hypervisor may allow a guest OS running in a virtual machine
to directly control a physical device that sends MSIs:
First, each virtual hart must have a guest interrupt file assigned to
it, giving each its own apparent IMSIC within the virtual machine.
Second, interrupts from the device must be signaled by wire through
an APLIC that can translate these interrupts into MSIs, or the
system must have an \mbox{IOMMU} that can translate the addresses of MSI
memory writes made by the device itself.

If a guest OS directly controls a device capable of sending MSIs, it
will naturally configure MSIs at the device with the guest physical
addresses the OS sees for the IMSICs of its virtual harts, not knowing
that these addresses are only virtual.
When the device performs a memory write for an MSI, the destination
address of this write must be translated by the \mbox{IOMMU} from the
guest physical address assigned by the guest OS into the true physical
address of the target guest interrupt file, using a translation table
supplied by the hypervisor.

By design, the translation an \mbox{IOMMU} must do for device MSIs is
fundamentally no different than the address translation the \mbox{IOMMU}
already must perform for other memory accesses from the same device,
converting guest physical addresses into true physical addresses.
Because each virtual hart is assigned a dedicated, physical guest
interrupt file that is indistinguishable from a true supervisor-level
interrupt file, no translation is needed for the data of an MSI write,
which specifies the interrupt's identity number in the target interrupt
file.

@@

\item
Modify the relevant translation tables at all \mbox{IOMMUs} so that MSIs for
this virtual interrupt file are now sent to the new physical interrupt
file.
Likewise, if any interrupts at an APLIC are forwarded by MSIs to the old
interrupt file, reconfigure the APLIC to send them to the new interrupt
file.
As needed, synchronize with all \mbox{IOMMUs} and APLICs to ensure that no
straggler MSIs will arrive at the old interrupt file after this step.
Synchronizing with an APLIC can be accomplished using the
algorithm of Section~\ref{sec:AdvPLIC-MSISync}.

### 2022-08-10 doc/src/intro.tex
Replace "I/O MMU" with "IOMMU"

\item
With the help of an I/O MMU (memory management unit) for redirecting
MSIs, maximize the opportunities and ability for a guest operating
system running in a virtual machine to have direct control of devices
with minimal involvement of a hypervisor.

@@

\begin{table*}[h!]
\begin{center}
\begin{tabular}{|l|c|l|}
\hline
   & Maximum      & \multicolumn{1}{c|}{Requirements} \\
\hline
\hline
Physical harts
   & 16,384       & \\
\hline
Active virtual harts having direct control of \quad
   & 31 for RV32, & {\RISCV} hypervisor extension; \\
\ a device, per physical hart
   & 63 for RV64  & \ IMSICs with guest interrupt\\
   &              & \ files; and an I/O MMU \\
\hline
Idle (swapped-out) virtual harts having \quad
   & potentially  & An I/O MMU with support \\
\ direct control of a device, per physical hart \quad
   & thousands    & \ for memory-resident \\
   &              & \ interrupt files \\
\hline
Wired interrupts at a single APLIC \quad
   & 1023         & \\
\hline
Distinct identities usable for MSIs at each \quad
   & 2047         & IMSICs \\
\ hart (physical or virtual)
   &              & \\
\hline
\end{tabular}
\end{center}
\caption{%
Absolute limits on the numbers of harts and interrupt identities in a
system.
Individual implementations are likely to have smaller limits.%
}
\label{tab:overallLimits}
\end{table*}

@@

If {\RISCV} harts implement the Privileged Architecture's hypervisor
extension, IMSICs may have additional \emph{guest interrupt files} for
delivering interrupts to virtual machines.
Besides Chapter~\ref{ch:IMSIC} on the IMSIC, see
Chapter~\ref{ch:VSLevel} which specifically covers interrupts to
virtual machines.
If the system also contains an I/O MMU to perform address translation
of memory accesses made by I/O devices, then MSIs from those same
devices may require special handling.
This topic is addressed in Chapter~\ref{ch:IOMMU}, ``I/O MMU Support
for MSIs to Virtual Machines.''

>>>>

\item
With the help of an \mbox{IOMMU} (I/O memory management unit) for redirecting
MSIs, maximize the opportunities and ability for a guest operating
system running in a virtual machine to have direct control of devices
with minimal involvement of a hypervisor.

@@

\begin{table*}[h!]
\begin{center}
\begin{tabular}{|l|c|l|}
\hline
   & Maximum      & \multicolumn{1}{c|}{Requirements} \\
\hline
\hline
Physical harts
   & 16,384       & \\
\hline
Active virtual harts having direct control of \quad
   & 31 for RV32, & {\RISCV} hypervisor extension; \\
\ a device, per physical hart
   & 63 for RV64  & \ IMSICs with guest interrupt\\
   &              & \ files; and an IOMMU \\
\hline
Idle (swapped-out) virtual harts having \quad
   & potentially  & An IOMMU with support \\
\ direct control of a device, per physical hart \quad
   & thousands    & \ for memory-resident \\
   &              & \ interrupt files \\
\hline
Wired interrupts at a single APLIC \quad
   & 1023         & \\
\hline
Distinct identities usable for MSIs at each \quad
   & 2047         & IMSICs \\
\ hart (physical or virtual)
   &              & \\
\hline
\end{tabular}
\end{center}
\caption{%
Absolute limits on the numbers of harts and interrupt identities in a
system.
Individual implementations are likely to have smaller limits.%
}
\label{tab:overallLimits}
\end{table*}

@@

If {\RISCV} harts implement the Privileged Architecture's hypervisor
extension, IMSICs may have additional \emph{guest interrupt files} for
delivering interrupts to virtual machines.
Besides Chapter~\ref{ch:IMSIC} on the IMSIC, see
Chapter~\ref{ch:VSLevel} which specifically covers interrupts to
virtual machines.
If the system also contains an \mbox{IOMMU} to perform address translation
of memory accesses made by I/O devices, then MSIs from those same
devices may require special handling.
This topic is addressed in Chapter~\ref{ch:IOMMU}, ``\mbox{IOMMU} Support
for MSIs to Virtual Machines.''
